---
title: "Prostate Cancer"
author: "Paul Hendricks"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Prostate}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Overview

In this vignette, we demonstrate the power of `easyml` using a Prostate Cancer dataset. 

## Load the data

First we load our libraries and the Prostate dataset. 

```{r}
library(easyml)
library(dplyr)
library(ggplot2)

data("prostate", package = "easyml")
knitr::kable(head(prostate))
```

## Train a support vector machine model

To run an `easy_support_vector_machine` model, we pass in the following parameters:

* the data set `prostate`,
* the name of the dependent variable e.g. `lpsa`,
* whether to run a `gaussian` or a `binomial` model, 
* how to preprocess the data; in this case, we use `preprocess_scale` to scale the data, 
* the random state, 
* whether to display a progress bar, 
* how many cores to run the analysis on in parallel.

```{r warning=FALSE}
results <- easy_support_vector_machine(prostate, "lpsa", 
                                       preprocess = preprocess_scale, 
                                       n_samples = 10, n_divisions = 10, 
                                       n_iterations = 2, random_state = 12345, 
                                       n_core = 1)
```

## Assess results

Now let's assess the results of the `easy_support_vector_machine` model.

### Predictions: ROC Curve

We can examine both the in-sample and out-of-sample ROC curve plots for one particular trian-test split determined by the random state and determine the Area Under the Curve (AUC) as a goodness of fit metric. Here, we see that the in-sample AUC is higher than the out-of-sample AUC, but that both metrics indicate the model fits relatively well.

```{r}
results$plot_predictions_train_mean
results$plot_predictions_test_mean
```

### Metrics: AUC

We can examine both the in-sample and out-of-sample AUC metrics for `n_divisions` train-test splits (ususally defaults to 1,000). Again, we see that the in-sample AUC is higher than the out-of-sample AUC, but that both metrics indicate the model fits relatively well.

```{r}
results$plot_metrics_train_mean
results$plot_metrics_test_mean
```

## Discuss

In this vignette we used `easyml` to easily build and evaluate a support vector machine model TO BE EDITED.
